\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}

Modern software depends heavily on open-source components, yet downstream
projects remain exposed when an upstream repository quietly stops being
maintained. This thesis asks to what extent publicly observable repository
health and maintenance indicators can predict whether an open-source dependency
becomes inactive within a twelve-month horizon, and how the resulting risk score
can support risk-based dependency triage. Following a Design Science Research
approach, the work develops and evaluates OSS Risk Radar, an artifact that
constructs a leakage-controlled historical dataset from GH Archive event data
and public package metadata, engineers observation-time activity, responsiveness,
and contributor-concentration features, and trains calibrated inactivity-risk
models.

On a repository foundation dataset of 5{,}000 projects and 50{,}000 quarterly
observation snapshots, a gradient-boosted model reaches a held-out ROC-AUC of
\(0.945\) with a Brier score of \(0.089\), clearly outperforming a trivial
recency baseline (ROC-AUC \(0.655\)). The skill largely persists under a
repository-disjoint split (\(0.940\)) and among repositories that are still
active at the observation time (\(0.883\)), indicating that the model detects
early decline rather than merely restating existing quietness. A human-actor
filter is validated by ablation, and security-practice signals are kept as
parallel triage context rather than as predictors of inactivity. A practical
demonstrator applies the deployed model per repository to the dependency
repositories of a real project, producing a calibrated, confidence-aware, and
evidence-backed ranked triage view.
The score is positioned as a decision-support signal for prioritizing dependency
review, not as a verdict on abandonment.

\chapter*{Declaration of Authorship}
\addcontentsline{toc}{chapter}{Declaration of Authorship}

I declare that I wrote this thesis independently and used only the sources and
tools listed in the bibliography and acknowledgements.

\vspace{2cm}
\begin{tabular}{p{6cm}p{6cm}}
  \hrulefill & \hrulefill \\
  Place, date & Signature
\end{tabular}
