\chapter{Results and Discussion}
\label{chap:evaluation}

This chapter reports and interprets the empirical results of the trained
artifacts: predictive performance against baselines, calibration behaviour,
feature attribution and error analysis, and an ablation of the human-actor
filter.

\section{Predictive Performance}
\label{sec:results-performance}

Table~\ref{tab:model-comparison} reports held-out test metrics for the three
learned model families in both feature regimes, together with the trivial recency
baseline. All learned models were trained on the same time-aware split
(\(75/15/10\); \(n_\text{test}=4{,}183\) labelled observations) and calibrated on
the validation slice with the same histogram calibrator, so the comparison
isolates the effect of model family rather than data or calibration differences.
The combined quality score summarizes ranking skill (ROC-AUC) and calibration
(Brier) into a single comparison figure and is not intended as a standalone proof.
All reported metrics are defined formally in
Section~\ref{sec:method-metric-definitions}.

\begin{table}[ht]
  \centering
  \caption{Held-out test metrics by model family and feature regime. Higher is
  better for ROC-AUC, F1, and quality; lower is better for Brier and ECE. The
  best value within each regime is shown in bold.}
  \label{tab:model-comparison}
  \begin{tabular}{lrrrrr}
    \toprule
    Model & ROC-AUC & Brier & ECE & F1 & Quality \\
    \midrule
    \texttt{logistic-regression-full-history} & 0.930 & 0.097 & \textbf{0.012} & 0.883 & 0.759 \\
    \texttt{xgboost-full-history}             & \textbf{0.945} & \textbf{0.089} & 0.016 & \textbf{0.891} & \textbf{0.790} \\
    \texttt{neural-net-full-history}          & 0.939 & 0.092 & 0.013 & 0.888 & 0.778 \\
    \midrule
    \texttt{logistic-regression-cold-start}   & 0.885 & 0.127 & 0.049 & 0.857 & 0.656 \\
    \texttt{xgboost-cold-start}               & \textbf{0.940} & \textbf{0.093} & \textbf{0.011} & \textbf{0.887} & \textbf{0.777} \\
    \texttt{neural-net-cold-start}            & 0.937 & 0.095 & \textbf{0.011} & 0.884 & 0.771 \\
    \midrule
    \texttt{recency-baseline} (days since commit) & 0.655 & --- & --- & 0.718 & --- \\
    \bottomrule
  \end{tabular}
\end{table}

All three families rank inactivity well above chance in both regimes. In the
full-history regime the gradient-boosted trees (\texttt{xgboost-full-history})
achieve the strongest ranking and calibration (ROC-AUC \(0.945\), Brier
\(0.089\)), ahead of the multilayer perceptron (\(0.939\) / \(0.092\)) and
logistic regression (\(0.930\) / \(0.097\)); logistic regression attains the
lowest expected calibration error but at a clear cost in ranking skill. The
cold-start regime preserves the same ordering, with the gap between the linear
model and the two non-linear models widening (logistic regression falls to
ROC-AUC \(0.885\) and Brier \(0.127\)), which is consistent with non-linear
models exploiting interactions among the reduced cold-start feature set.

The neural network was included as a deliberate baseline to test whether a
higher-capacity non-linear model is warranted for this tabular feature space. It
does not improve on the gradient-boosted trees on any reported metric in either
regime, while adding training cost, weaker interpretability, and additional
calibration sensitivity. This is consistent with the broader finding that
tree-based models remain a strong default on moderate-sized tabular
problems~\cite{grinsztajn2022trees}. Accordingly, logistic regression is
retained for its transparent coefficient attribution and XGBoost as the deployed
runtime scorer, while the neural network is reported as a tested-and-rejected
alternative rather than promoted into the scoring ensemble.

Two runtime compositions of these evaluated models are reported as engineering
mechanisms rather than as separately benchmarked results. At scoring time the
backend averages the staged models of a regime into an ensemble profile and
attaches a model-agreement factor, and it derives a per-prediction confidence
from signal completeness and prediction decisiveness. Both are exercised in the
demonstrator (Chapter~\ref{chap:demonstrator}); the headline metrics in this
chapter are, however, always reported for the individual artifacts, so no claim
is made that the runtime ensemble outperforms its strongest member. Quantifying a
possible ensemble gain is left to future work (Chapter~\ref{chap:conclusion}).

The learned models are worth reporting only if they improve on the trivial
recency signal that the label is partly built from
(Section~\ref{sec:method-model-training}). The last row of
Table~\ref{tab:model-comparison} reports this baseline: ranking the held-out test
slice purely by days since the last commit attains a ROC-AUC of only \(0.655\),
and selecting an F1-optimal single threshold on the validation slice degenerates
to predicting inactivity for essentially every repository (threshold at zero days,
recall \(1.0\), F1 \(0.718\) at the balanced base rate). Every learned model
clears this bar by a wide margin --- the deployed \texttt{xgboost-full-history}
model raises ROC-AUC from \(0.655\) to \(0.945\) --- so the models add substantial
ranking skill over recency alone rather than merely restating it. To further
separate genuine predictive skill from activity persistence, the deployed
full-history model is additionally re-scored on the active-at-\(t\) subset defined
in Section~\ref{sec:method-evaluation-strategy}; this slice analysis is reported in
Section~\ref{sec:results-features}, where the same recency baseline is also
revisited.

\section{Calibration and Threshold Selection}
\label{sec:results-calibration}

Because the score is intended to be read as a probability rather than only as a
ranking, calibration is evaluated alongside discrimination. Every model's raw
outputs are mapped through a histogram calibrator fitted on the validation slice
and never on the test slice, so the reported calibration reflects held-out
behaviour. For the deployed \texttt{xgboost-full-history} model this yields a
Brier score of \(0.089\), a log loss of \(0.284\), and an expected calibration
error of \(0.016\) at the default decision threshold of \(0.5\), where the
operating point attains precision \(0.87\) and recall \(0.92\). The linear
\texttt{logistic-regression-full-history} model reaches the lowest expected
calibration error (\(0.012\)) but a weaker Brier score (\(0.097\)), consistent
with its lower ranking skill in Table~\ref{tab:model-comparison}.

The reliability of the calibrated probabilities is well behaved: the empirical
inactivity rate rises monotonically across the calibration bins, from
approximately \(0.02\) in the lowest predicted-probability band to \(0.98\) in
the highest, so higher scores do correspond to higher observed inactivity
frequencies rather than only to a higher rank. Two caveats apply. First, the two
extreme bands hold most of the held-out mass, so mid-range probabilities are
supported by fewer samples and are correspondingly less certain; the
per-prediction confidence surfaced by the demonstrator (Chapter~\ref{chap:demonstrator})
reflects exactly this bin-level support. Second, calibration is conditional on
the sampled prevalence rather than a natural deployment prevalence
(Section~\ref{sec:method-limitations}), so the absolute probabilities would
require recalibration before being read as live population probabilities.

\section{Feature Importance and Error Analysis}
\label{sec:results-features}

\subsection{Feature Attribution}
\label{sec:results-attribution}

Feature attribution is read from the logistic-regression model, which is
retained precisely for its transparent coefficients (Section~\ref{sec:results-performance}).
Because the inputs are standardized, the coefficient magnitudes are directly
comparable, and their signs indicate the direction of the association with
inactivity. The strongest inactivity-\emph{lowering} signals are the breadth of
sustained development (\texttt{active\_commit\_months\_365d}, standardized
coefficient \(-0.94\)) and recent maintainer integration
(\texttt{has\_recent\_pr\_merge\_flag}, \(-0.47\)); the strongest inactivity-\emph{raising}
signals are staleness and slow responsiveness, including days since the last
commit and last push (each \(\approx +0.19\)), a slower median issue first
response (\(+0.16\)), and a larger stale-issue share (\(+0.15\)). These
directions match the intended semantics of the features
(Appendix~\ref{appendix:feature-reference}). A few coefficients are less
intuitive (for example a positive association for the log-scaled star count),
which reflects partial associations under the sampled prevalence and
correlations among features rather than a causal effect; the coefficients are
therefore used to explain \emph{direction} for a given repository, while ranking
performance is carried by the gradient-boosted model.

The gradient-boosted model exposes a complementary, non-linear view through its
gain-based feature importances (normalized to sum to one over the 39 features
that receive at least one split). Its ranking is dominated by recent commit
volume: \texttt{commits\_365d} (\(0.29\)) and \texttt{commits\_90d} (\(0.26\))
together account for over half of the total gain, followed by
\texttt{recent\_contributors\_90d} (\(0.12\)) and
\texttt{active\_commit\_months\_365d} (\(0.05\)); pull-request and issue-flow
features (\texttt{opened\_prs\_90d}, \texttt{open\_issue\_growth\_90d},
\texttt{merged\_prs\_90d}), popularity (\texttt{stars\_log1p}), and release
activity (\texttt{releases\_365d}) contribute the next tier. The two views agree
on the substantive story --- sustained recent development activity and
contributor breadth are the strongest signals --- while differing in emphasis:
the linear model leans on the breadth of active commit months, whereas the
gradient-boosted model concentrates gain on the raw recent-commit-volume windows.

\subsection{Robustness to Repository Recurrence}
\label{sec:results-robustness}

Under the repository-disjoint split (Section~\ref{sec:impl-grouped-split}),
where every snapshot of a repository is confined to one partition, held-out
ROC-AUC decreases only from \(0.945\) to \(0.940\) and the Brier score rises only
from \(0.089\) to \(0.090\). The small gap bounds the optimism attributable to a
repository appearing in both training and test slices and indicates that the
discrimination generalizes to unseen repositories rather than resting on
repository memorization.

\subsection{Slice Evaluation}
\label{sec:results-slices}

A single aggregate metric can average a strong subgroup with a weak one. The
held-out predictions of the deployed \texttt{xgboost-full-history} model are
therefore re-scored within subgroups that vary in ways relevant to deployment
(Table~\ref{tab:slice-metrics}).

\begin{table}[ht]
  \centering
  \caption{Held-out ROC-AUC and Brier score of \texttt{xgboost-full-history}
  within subgroups of the test slice. ``Inactive rate'' is the positive rate in
  the subgroup.}
  \label{tab:slice-metrics}
  \begin{tabular}{llrrr}
    \toprule
    Dimension & Subgroup & \(n\) & Inactive rate & ROC-AUC \\
    \midrule
    Popularity      & high            & 2178 & 0.48 & 0.968 \\
    Popularity      & medium          & 1398 & 0.66 & 0.945 \\
    Popularity      & low             & 607  & 0.64 & 0.689 \\
    \midrule
    Activity at \(t\) & active@\(t\)   & 1786 & 0.17 & 0.883 \\
    Activity at \(t\) & quiet@\(t\)    & 2397 & 0.85 & 0.860 \\
    \midrule
    History regime  & has-history     & 2320 & 0.30 & 0.917 \\
    History regime  & cold-start-like & 1863 & 0.88 & 0.872 \\
    \bottomrule
  \end{tabular}
\end{table}

Three findings follow. First, the aggregate ROC-AUC masks a much weaker
obscure-repository subgroup: low-popularity repositories are ranked at only
\(0.689\), against \(0.968\) for high-popularity ones, so scores for
low-visibility projects deserve more caution and are exactly where the
per-prediction confidence should be low. Second, and most importantly for the
activity-persistence threat (Section~\ref{sec:validity-internal}), the model
still separates outcomes well \emph{among repositories that are active at the
observation time} (ROC-AUC \(0.883\) on the active-at-\(t\) subset, where the
inactive rate is only \(0.17\)); the headline metric is therefore not merely a
restatement of ``already-quiet repositories stay quiet''. On this same
active-at-\(t\) subset the trivial recency baseline reaches only ROC-AUC
\(0.777\), so the learned model retains a clear margin over recency exactly where
early-warning skill matters most. Third, full history
helps: repositories with reconstructable history are ranked at \(0.917\) versus
\(0.872\) for cold-start-like repositories, consistent with the two-regime
design (Section~\ref{sec:method-prediction-regimes}).

\subsection{Error Analysis}
\label{sec:results-error-analysis}

The comparison against an explicit trivial recency baseline is now part of the
reported results (Section~\ref{sec:results-performance} and the slice analysis
above); the systematic formative case review that motivated the feature-set
revision (Section~\ref{sec:method-iterative-development}) remains a complementary
qualitative analysis and is noted as a remaining item in
Chapter~\ref{chap:conclusion}. The slice results above already give the main
error-analysis conclusion: the dominant residual weakness is low-visibility
repositories with sparse public signals, which the confidence layer is designed
to flag rather than hide.

\section{Ablation: The Human-Actor Filter}
\label{sec:results-bot-ablation}

Commit- and contributor-based features and the maintenance label are computed
from human actors only; automated (bot) accounts are excluded
(Section~\ref{sec:method-feature-engineering}). Because this is a design choice
that could plausibly be omitted, it is tested directly by an ablation: the full
5{,}000-repository dataset was rebuilt with the human-actor filter disabled
through a single build flag, with every other setting identical, yielding
50{,}000 snapshots that align one-to-one with the filtered build. Three effects
are examined.

\textbf{Label corruption.} Disabling the filter flips 357 of 50{,}000 labelled
snapshots (\(0.71\%\)) from inactive to active, and \emph{none} in the opposite
direction. The effect is strictly one-directional: a bot commit or merge inside
the future window can satisfy the two-of-four maintenance rule, so an otherwise
dormant repository is relabelled as maintained, whereas bots never make an
active repository look inactive. Omitting the filter would therefore bias the
target by systematically \emph{masking} inactivity, which is the error direction
that matters most for a risk signal.

\textbf{Feature inflation.} On the observation-time side, bots inflate the
activity features substantially: with the filter disabled, \texttt{commits\_365d}
changes on 7{,}858 of 50{,}000 snapshots (\(15.7\%\)) by a mean of \(+938\)
commits, \texttt{commits\_90d} by \(+388\), and \texttt{contributors\_365d} by
\(+1.4\) distinct contributors, while pull-request \emph{count} features are
unaffected because they are not actor-filtered. A small number of bot-driven
repositories would thus dominate the activity signal.

\textbf{Predictive effect.} Retraining all four model families on the unfiltered
dataset changes the held-out metrics only at the noise level
(\(|\Delta\text{ROC-AUC}| \le 0.003\), \(|\Delta\text{Brier}| \le 0.002\));
measured against each dataset's own labels, bots appear harmless. This is
expected but misleading, because the unfiltered model is then scored against a
partially corrupted target. A cleaner test trains on the bot-contaminated
\emph{features} while evaluating against the clean, human-filtered \emph{labels}
(Table~\ref{tab:bot-ablation-crosstest}): the contaminated features are
consistently worse across all four families, though by a small margin (ROC-AUC
falls by \(0.0007\)--\(0.0018\); Brier rises by up to \(0.0018\)).

\begin{table}[ht]
  \centering
  \caption{Cross-test of the human-actor filter: training on bot-contaminated
  features versus clean features, both evaluated against the identical clean
  (human-filtered) label on the held-out slice. A lower ROC-AUC or higher Brier
  for the contaminated features means the filter helps.}
  \label{tab:bot-ablation-crosstest}
  \begin{tabular}{lrrrr}
    \toprule
    Model & ROC-AUC (clean) & ROC-AUC (contam.) & \(\Delta\)ROC-AUC & \(\Delta\)Brier \\
    \midrule
    \texttt{logistic-regression-full-history} & 0.9379 & 0.9373 & \(-0.0007\) & \(+0.0009\) \\
    \texttt{xgboost-full-history}             & 0.9529 & 0.9522 & \(-0.0007\) & \(+0.0001\) \\
    \texttt{logistic-regression-cold-start}   & 0.8889 & 0.8871 & \(-0.0018\) & \(+0.0018\) \\
    \texttt{xgboost-cold-start}               & 0.9488 & 0.9477 & \(-0.0011\) & \(+0.0018\) \\
    \bottomrule
  \end{tabular}
\end{table}

\textbf{Interpretation.} The filter's justification is construct validity rather
than discrimination. Its predictive cost when omitted is small, since the
gradient-boosted models are largely robust to the minority of bot-inflated
repositories, but it is consistently in the expected direction, and, more
importantly, the filter removes a one-directional label bias that a headline
metric computed on the corrupted data does not reveal. The human-actor filter is
therefore retained. Absolute metrics in this ablation differ slightly from
Table~\ref{tab:model-comparison} because the ablation uses a freshly rebuilt,
fully labelled dataset; the comparison is internally controlled and varies only
the filter.
