\chapter{Practical Demonstrator}
\label{chap:demonstrator}

This chapter demonstrates how the trained inactivity-risk model is applied end to
end, from a set of dependency repositories to a ranked risk view that supports
triage. While Chapter~\ref{chap:evaluation} evaluates the model on the held-out
test set, this chapter shows the artifact operating on a realistic input, as
required by the demonstration phase of the Design Science Research process
\cite{hevner2004design,peffers2007design}. The demonstrator is run against the
deployed OSS Risk Radar instance using the staged model bundle (primary
full-history artifact \texttt{xgboost-full-history}, version \texttt{0.2.0});
all figures reported here are real outputs of that deployed artifact and reflect
the public repository state at the time of analysis.

\section{Example Repository Set}
\label{sec:demo-mapping}

The unit of analysis is a single repository, and each repository is scored
independently. Enumerating which repositories a project depends on is the task of
established software-composition-analysis tooling, such as the OSS Review Toolkit
\cite{ortAnalyzer}; the artifact treats this dependency inventory as an external
input rather than resolving dependency graphs itself.

For this demonstration the input is the set of dependency repositories of
\texttt{ArchipelagoMW/Archipelago}, a large and actively developed open-source
multi-world game randomizer. Its declared Python dependencies correspond to 30
source repositories, ranging from widely used infrastructure libraries (for
example \texttt{flask}, \texttt{jinja2}, \texttt{werkzeug}) to smaller, more
specialized packages. This project was chosen because that mixture is expected to
produce a non-trivial triage view rather than a uniformly low-risk one.

Each repository is scored through the deployed analysis workflow, which enriches
live repository metadata and applies the staged model ensemble. The two prediction
regimes of Section~\ref{sec:method-prediction-regimes} appear directly in this
run: one repository (\texttt{flask}) is present in the staged repository feature
cache and is therefore scored in the richer full-history regime, while the
remaining 29 are absent from the foundation cache and fall back to the cold-start
regime. This distribution is itself an honest illustration of the cold-start
coverage limitation highlighted in Chapter~\ref{chap:conclusion}: an arbitrary
project's dependency repositories are mostly outside the precomputed foundation
seed and are consequently scored by the weaker cold-start model, a fact the
confidence layer is designed to surface rather than hide.

\section{Risk Ranking Output and Interpretation Guidance}
\label{sec:demo-ranking}

Table~\ref{tab:demo-ranking} shows the ranked inactivity-risk view produced by the
scoring service for the 30 repositories. For each repository the output includes
the inactivity-risk score (0--100), the discrete action level (\texttt{monitor},
\texttt{review}, or \texttt{replace\_candidate}), a confidence value derived from
signal completeness and prediction decisiveness, and the prediction regime. Of the
30 scored repositories, the distribution is one high-risk item (action
\texttt{review}), five medium-risk items, and 24 low-risk items; all but one are
assigned the \texttt{monitor} action, so the ranking concentrates analyst
attention on a single repository rather than diluting it across the whole set.

\begin{table}[htbp]
  \centering
  \small
  \caption{Ranked inactivity-risk view for the dependency repositories of
  \texttt{ArchipelagoMW/Archipelago}, as produced by the deployed model. Risk is
  the 0--100 inactivity-risk score; ``Conf.'' is the per-prediction confidence;
  ``Regime'' is full-history (FH) or cold-start (CS).}
  \label{tab:demo-ranking}
  \begin{tabular}{llrlrl}
    \toprule
    Package & Source repository & Risk & Action & Conf. & Regime \\
    \midrule
    bsdiff4           & \texttt{ilanschnell/bsdiff4}        & 77.9 & review  & 0.69 & CS \\
    \midrule
    cymem             & \texttt{explosion/cymem}            & 46.6 & monitor & 0.81 & CS \\
    kivymd            & \texttt{kivymd/KivyMD}              & 46.6 & monitor & 0.71 & CS \\
    pymem             & \texttt{srounet/Pymem}              & 46.6 & monitor & 0.71 & CS \\
    markupsafe        & \texttt{pallets/markupsafe}         & 46.6 & monitor & 0.81 & CS \\
    setproctitle      & \texttt{dvarrazzo/py-setproctitle}  & 46.6 & monitor & 0.67 & CS \\
    \midrule
    jinja2            & \texttt{pallets/jinja}              & 26.0 & monitor & 0.70 & CS \\
    schema            & \texttt{keleshev/schema}            &  8.8 & monitor & 0.86 & CS \\
    platformdirs      & \texttt{tox-dev/platformdirs}       &  8.8 & monitor & 0.91 & CS \\
    typing-extensions & \texttt{python/typing\_extensions}  &  8.8 & monitor & 0.86 & CS \\
    pyshortcuts       & \texttt{newville/pyshortcuts}       &  8.8 & monitor & 0.86 & CS \\
    pony              & \texttt{ponyorm/pony}               &  8.8 & monitor & 0.86 & CS \\
    flask-caching     & \texttt{pallets-eco/flask-caching}  &  8.8 & monitor & 0.91 & CS \\
    colorama          & \texttt{tartley/colorama}           &  5.8 & monitor & 0.83 & CS \\
    kivy              & \texttt{kivy/kivy}                  &  5.8 & monitor & 0.94 & CS \\
    certifi           & \texttt{certifi/python-certifi}     &  5.8 & monitor & 0.78 & CS \\
    orjson            & \texttt{ijl/orjson}                 &  5.8 & monitor & 0.89 & CS \\
    pathspec          & \texttt{cpburnz/python-pathspec}    &  5.8 & monitor & 0.89 & CS \\
    werkzeug          & \texttt{pallets/werkzeug}           &  5.8 & monitor & 0.89 & CS \\
    waitress          & \texttt{Pylons/waitress}            &  5.8 & monitor & 0.94 & CS \\
    flask-compress    & \texttt{colour-science/flask-compress} & 5.8 & monitor & 0.78 & CS \\
    flask-limiter     & \texttt{alisaifee/flask-limiter}    &  5.8 & monitor & 0.89 & CS \\
    flask-cors        & \texttt{corydolphin/flask-cors}     &  5.8 & monitor & 0.89 & CS \\
    websockets        & \texttt{python-websockets/websockets} & 1.8 & monitor & 0.98 & CS \\
    pyyaml            & \texttt{yaml/pyyaml}                &  1.8 & monitor & 0.93 & CS \\
    cython            & \texttt{cython/cython}              &  1.8 & monitor & 0.98 & CS \\
    bokeh             & \texttt{bokeh/bokeh}                &  1.8 & monitor & 0.87 & CS \\
    mistune           & \texttt{lepture/mistune}            &  1.8 & monitor & 0.93 & CS \\
    psutil            & \texttt{giampaolo/psutil}           &  1.8 & monitor & 0.81 & CS \\
    flask             & \texttt{pallets/flask}              &  1.7 & monitor & 0.97 & FH \\
    \bottomrule
  \end{tabular}
\end{table}

\paragraph{The flagged repository.} The single item that the ranking elevates for
review is \texttt{bsdiff4} (\texttt{ilanschnell/bsdiff4}), a small binary-diff
library with a risk score of \(77.9\). Its explanation factors, taken directly
from the scored output, are consistent with the intended semantics of the model:
the two cold-start ensemble members predict a \(78.7\%\) and \(77.1\%\) inactivity
risk respectively, the model-agreement factor reports a narrow spread between
them, and the input-completeness factor records that \(88\%\) of expected signals
were available before imputation. The score reflects a genuinely quiet
maintenance profile (a low-popularity repository with no contributor activity in
the recent window and no detected release cadence), which is exactly the kind of
small, single-maintainer dependency that a supply-chain reviewer would want
surfaced. The confidence of \(0.69\) is moderate rather than high, and a caveat
records that the score was produced by the cold-start model because no
full-history cache row was available, so the flag is presented as a prompt for
manual review rather than as a definitive judgement.

\paragraph{A low-risk repository for contrast.} At the opposite end,
\texttt{flask} (\texttt{pallets/flask}) receives a risk score of \(1.7\) with a
confidence of \(0.97\). It is the one repository scored in the full-history
regime, and both full-history ensemble members agree on a \(1.7\%\) inactivity
risk with \(98\%\) signal completeness. This contrast between a moderate-confidence
cold-start flag and a high-confidence full-history clearance illustrates why the
confidence value, not only the risk score, is part of the triage output.

Interpretation guidance follows the rest of the thesis. The score is a
prioritization signal, not a trust or security verdict: a low score does not
certify a repository as secure, and the elevated score for \texttt{bsdiff4}
indicates thin public maintenance signals rather than a known defect. Low
confidence or a large number of imputed signals should reduce the decisiveness of
any action, and cold-start scores in particular are conditional on the reduced
feature set and on the sampled prevalence to which the model is calibrated
(Section~\ref{sec:method-limitations}), so they warrant recalibration before being
read as absolute deployment probabilities. Used this way, the ranked view converts
a flat list of 30 repositories into a single reviewable candidate plus a small
medium-risk cluster, which is the intended practical contribution of the artifact.
