\chapter{Conclusion and Outlook}
\label{chap:conclusion}

This thesis asked to what extent publicly observable repository health and maintenance indicators can predict 12-month OSS dependency inactivity, and how the resulting inactivity-risk score can be combined with parallel security-practice signals to support dependency triage. The empirical results support a qualified positive answer. On a time-aware held-out split, the deployed full-history gradient-boosted model ranks inactivity with a ROC-AUC of \(0.945\) and is well calibrated (Brier \(0.089\), expected calibration error \(0.016\)). The skill largely persists under a repository-disjoint split (\(0.940\)), so it generalizes to unseen repositories rather than memorizing them, and it remains substantial among repositories that are active at the observation time (\(0.883\)), so it is not merely a restatement of activity persistence.

At the same time, the model is markedly weaker on low-visibility repositories (ROC-AUC \(0.689\)) and in the cold-start regime, and its probabilities are calibrated to a sampled rather than a natural prevalence. The score is therefore presented as a calibrated triage aid with an explicit per-prediction confidence that flags exactly these weak cases, not as a verdict on abandonment. Consistent with the research design, security-practice signals are surfaced as complementary triage context and are deliberately excluded from the inactivity model, so the thesis does not claim that they improve inactivity prediction. Applied per repository to the dependency repositories of a project, whose inventory is supplied by an external software-composition-analysis tool such as the OSS Review Toolkit \cite{ortAnalyzer}, these scores produce the ranked, evidence-backed triage view demonstrated in Chapter~\ref{chap:demonstrator}, which is the intended practical contribution.

\section{Outlook and Future Work}
\label{sec:conclusion-outlook}

Several extensions are deliberately scoped out of the completed method and are recorded here as future work rather than as claimed results.

\begin{itemize}
    \item \textbf{External validation.} The reported evaluation uses a time-aware held-out split within the constructed foundation dataset (Section~\ref{sec:method-evaluation-strategy}). A stronger robustness check would re-sample a fresh set of repositories that were not part of the foundation seed and evaluate the trained artifacts on that independent set. This would test generalization beyond the sampled prevalence and seed strata.
    \item \textbf{Systematic qualitative error analysis.} The completed artifacts cover Logistic Regression, XGBoost, and a feedforward neural network; the active-at-\(t\) slice isolates predictive skill from activity persistence (Section~\ref{sec:results-slices}), and the trivial recency baseline is now reported and clearly outperformed (Section~\ref{sec:results-performance}). A systematic, larger-scale qualitative case review --- beyond the formative examples that motivated the feature-set revision --- remains to be added as complementary evidence.
    \item \textbf{Additional model families.} Beyond the covered families, further extensions such as more expressive deep-learning architectures or sequence models over the repository event history are left as future work.
    \item \textbf{Ensemble evaluation.} The runtime averages the staged models of a regime into an ensemble profile (Section~\ref{sec:results-performance}), but the reported metrics are for the individual artifacts. Benchmarking whether the ensemble improves on its strongest member, and whether the model-agreement spread is a useful uncertainty signal, is left as future work.
    \item \textbf{Prevalence-aware calibration.} Because the seed oversamples dormant and archived repositories, predicted probabilities are calibrated to the sampled prevalence (Section~\ref{sec:method-limitations}). Recalibrating to an estimated deployment prevalence would make the scores directly usable as live triage probabilities.
    \item \textbf{Label sensitivity analysis.} The maintenance label uses heuristic thresholds (Section~\ref{sec:method-definitions}). A systematic sensitivity analysis over these thresholds would quantify how strongly the results depend on the specific label definition.
    \item \textbf{Cold-start coverage and on-demand history reconstruction.} Repositories outside the foundation seed are scored by the cold-start model, which uses only features derivable from a single live snapshot and therefore omits the strongest historical signals (most notably the distribution of commit activity across the year). Two observations make this a priority for future work. First, the held-out cold-start metrics are measured on seed-distribution repositories but applied to arbitrary submissions, so the deployed cold-start performance is likely optimistic; the per-prediction confidence already flags such out-of-distribution cases rather than hiding them. Second, a submitted repository's recent history can in principle be reconstructed on demand from the GitHub REST API (for example, weekly commit activity and per-contributor statistics) to recover several full-history features. This was deliberately not implemented, because the live API cannot supply the full feature set: package-registry features (package age, dependency count, published versions, popularity tier) are not exposed by the repository API, per-issue first-response latency requires prohibitively many timeline requests, and the year-over-year change features require a second year of history. Imputing this missing tail into the full-history model would itself introduce a distribution shift. A cleaner extension is therefore a third \emph{warm-start} model trained on exactly the subset of features that can be reconstructed reliably from the live API, validated for parity against the precomputed seed feature cache and supported by a rate-limit-aware request cache. This is left as future work because of the additional engineering and validation effort it requires.
\end{itemize}
